### 2020-04-28 submission/manuscript.tex
First submission

\vspace{25mm}

Ada K. Hagan\({^2}\), Nick Lesniak, \emph{Author names here, order
alphabetical}, Patrick D. Schloss\({^1}\)\({^\dagger}\)

\vspace{40mm}

@@

\(1\) Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

\(2\) Alliance SciComm \& Consulting, Linden, MI 48418

\vspace{35mm}

@@

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing data
generation and data analysis, we have found scientists feel empowered
when they can both ask and answer their own biological questions. Yet,
the standard undergraduate and graduate training in the biological
sciences is insufficient for developing these advanced data analysis
skills. In our experience performing microbiome research, it is more
common to find exceptional bench scientists who are inexperienced at
analyzing large data sets than to find the reverse. Of course this
raises a challenge: How do we train bench scientists that analyze data
sets to answer biological questions?

The ever-growing ability to generate data and constant feeling of
helplessness in analyzing it is analogous to the struggles we also face
with engaging the voluminous scientific literature. A common strategy to
keeping up with the literature are journal clubs, which involve group
discussion of a pre-selected paper. These range from informal
discussions to PowerPoint presentations and course credit. In addition
to staying current on the literature, journal clubs help strengthen
skills in critical thinking, communication, and literature review
{[}1{]}. Our research group has leveraged the similarity between the
overwhelming natures of scientific literature and data analysis to
address the challenge of teaching reproducible data analysis practices
to bench scientists. Over the past 4 years we have experimented with a
Code Club model to improve data analysis skills in a community
environment.

Our Code Club sessions are an hour long and alternate with Journal Club
as the second part of weekly two hour lab meetings. Presenters volunteer
to lead a session with the expectation that they alternate leading
Journal and Code Clubs. Initially, the Code Club was used to review code
from trainee projects. Instead of a presentation, the presenter would
project their code onto a screen and the participants would go line by
line through the code, stating the logic behind each line. This approach
emphasized the importance of code readability and gave beginners the
opportunity to see the real-life, messy code of more experienced peers.
Unfortunately, the format only allowed us to review a fraction of a
project's code base, making it difficult to integrate the programmer's
logic across their full project. A major issue with this model was that
sometimes beginners couldn't contribute to improving the code, and even
when they could, more senior group members would eventually dominate the
discussion. This led to a lack of participation by beginners, who would
mentally check out to focus on other activities, and frequently resulted
in an adversarial environment between the presenter and senior group
members. As a result, presenters rarely offered to present Code Club
again.

From these experiences, we began a collective conversation to improve
our Code Club model and have evolved it into two approaches. The first
is a more constructive version of a group code review. The presenter
clearly states the problem they want to solve, breaks the participants
into smaller groups, and then asks each group to solve the problem, or a
portion of it. For example, someone may have an R script with a
repeating chunk of code. The challenge for the session would be to
convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself){[}2{]}. The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, a more memorable activity was giving participants
raw data and a finished plot. Paired participants were tasked with
generating the plot using either base R or ggplot, but base R users had
to use ggplot and vice versa. In either approach, the Code Club ends
with a report back to the presenter and larger group describing the
approach each pair took. Our Code Clubs typically have 7 to 10
participants, but the inherent ``think-pair-share'' strategy should
allow it to be scaled to larger groups {[}3{]}.

We continue to experiment with approaches to running Code Club. During
the period of the Covid-19 pandemic in 2020 when many research labs were
shuttered, PDS started hosting weekly Code Club virtually. Participants
could engage the material posted to
\url{https://www.riffomonas.org/code_club/} via a Zoom meeting or
asynchronously using the recorded sessions available on YouTube.
Regardless of our approach to leading a Code Club, we have learned that
it is critical for the presenter to clearly articulate their goals and
facilitate participant engagement. Although any individual Code Club
session may be more experimental than another, on the whole they are a
critical tool to train bench scientists in reproducible data analysis
practices. We have provided some examples of successful code club topics
in Table 1. We have summarized the results of our own experimentation as
Code Club presenters and participants into Ten Rules. The first 3 rules
apply to all participants at every Code Club, while Rules 4 through 8
are targeted to Code Club presenters and the last two for participants.

@@

It is critical that the presenter and participants respect each other
and that a designated individual (e.g.~the lab director) enforces a code
of conduct (see X for an example). Each member of the Code Club must
have the humility to acknowledge that they have more to learn about any
given topic. We have found that many problems are avoided when the
presenter takes charge of the session with a clear lesson plan,
thoughtfully creating groups, and giving encouragement. Similarly,
participants foster a positive environment by remembering that the task
is not a competition, focusing on the presenter's goals, allowing their
partner to contribute, asking clarifying questions when appropriate, and
avoiding distractions (e.g.~email, social media). Learning to program is
challenging and too often results in a toxic environment of nitpicking
and nonconstructive criticism. All parties in a Code Club are
responsible for preventing this toxicity by demonstrating respect for
themselves and their colleagues.

@@

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it often led to frustration for both
the presenter and participants. If the presenter shared their own code,
did they want participants to focus on their coding style or did they
want help incorporating a new package into their workflow? Participants
will always notice or ask about code concepts that are not the focus of
the exercise; a presenter with a specific goal can bring tangential
conversations back to the planned task. Where possible, presenters
should create a simplified scenario (i.e.~minimal, reproducible example
{[}4{]}), which can be helpful in focusing the participants. The
presenter should verify ahead of time that the simplified example works
and behaves the way they expect. Beyond the content, clear goals for
participant activities will help both parties stay on task and avoid
frustration. For more advanced learners, the presenter can create
stretch goals or give an activity with multiple stopping points where
participants would feel successful (e.g.~commenting code, creating
function, implementing function, refactoring function). Accomplishing
specific goals is more likely to result in a positive outcome for
presenters and participants.

@@

Participants want to learn topics that will either be useful to them or
help their colleague (i.e., the presenter). Presenters should do their
best to satisfy those motivations, whether it is through the relevancy
of the concept or the data. It does not make sense to present a Code
Club on downloading stock market data if it is not useful or interesting
to the group. Similarly, participants should not be tasked with
improving the presenter's code if the presenter has no intention of
incorporating the suggestions. A list of packages or tasks that group
members are interested in could help an uninspired presenter choose a
topic that will make a rewarding Code Club.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, such as follows a turnover in group
membership. In these circumstances it can be useful to cover commonly
used tools but that might result in disinterest of more experienced
participants. We have identified several strategies to overcome the
challenges presented by participants at varying skill levels. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on their goals. Partnerships between those with similar skill
levels requires the presenter to design appropriate activities each
skill level. We have found that commenting code is a good skill for
beginners since it forces them to dissect and understand a code chunk
line-by-line. An advantage of partnerships with disparate skill levels
is assigning a single task, which provides the presenter with a diverse
range of methods that achieve the same result. This approach to pairing
also helps to graft new members into the group that have emerging
programming skills. Regardless of how partners are selected, consider
asking the pairs to identify a navigator and a driver {[}5{]}. The
driver types at the computer while the navigator tells them what to
type, thus ensuring participation of both partners. Midway through the
activity, the presenter can have the partners switch roles.
Intentionally forming pairs can also engineer group interactions by
breaking up cliques, avoiding potentially disruptive partnerships, or
pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion
about a paper that the participant has not read. In that context, coming
to Code Club without having installed a dependency is similar to asking
a simple question about the Journal Club paper. Both instances show a
lack of respect and preparation. Just as a presenter must follow Rule 6
to provide materials ahead of time, participants must review the code in
advance, download the data sets, install the necessary packages, and
perhaps read up on the topic. If the Code Club is based on a paper or
chapters in a data science book (e.g., {[}6--8{]}), the participants
should read them before the session and consider how they might
incorporate the concepts into their own work.

@@

An essential ingredient of any Code Club is active participation from
all parties. Having an open laptop on the table and permission to use it
can feel like an invitation to get distracted by other work, emails, and
browsing the internet; fight that urge and focus on the presenter's
goals. Be respectful and allow your partner to contribute. Speak up for
yourself and force your partner to let you contribute. If the material
seems too advanced for you, it can be frustrating, and tempting to
mentally check out. Fortunately, programming languages like R and Python
are generally expressive, which should allow you to engage with the
logic, even if the syntax is too advanced. Frankly, understanding the
logic of \emph{when} to use one modeling approach over another is more
important than knowing \emph{how} to use it. If you understand the
``why'', the ``how'' will quickly follow. More experienced participants
should aim to communicate feedback and coding suggestions at a level
that all participants can understand and engage in. Regardless of skill
levels, your partner and the presenter put themselves in a vulnerable
position by revealing what they do or do not know. Encourage them, and
show your gratitude for helping you learn something new, by fully
participating in each Code Club.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Lonsdale2016}{}%
1. Lonsdale A, Penington JS, Rice T, Walker M, Dashnow H. Ten simple
rules for a bioinformatics journal club. Lewitter F, editor. PLOS
Computational Biology. 2016;12: e1004526.
doi:\href{https://doi.org/10.1371/journal.pcbi.1004526}{10.1371/journal.pcbi.1004526}

\leavevmode\hypertarget{ref-Wilson2014}{}%
2. Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et al.
Best practices for scientific computing. Eisen JA, editor. PLoS Biology.
2014;12: e1001745.
doi:\href{https://doi.org/10.1371/journal.pbio.1001745}{10.1371/journal.pbio.1001745}

\leavevmode\hypertarget{ref-Kothiyal2014}{}%
3. Kothiyal A, Murthy S, Iyer S. Think-pair-share in a large CS1 class.
Proceedings of the 2014 conference on innovation \& technology in
computer science education - ITiCSE 14. ACM Press; 2014.
doi:\href{https://doi.org/10.1145/2591708.2591739}{10.1145/2591708.2591739}

\leavevmode\hypertarget{ref-StackOverflow_MREE}{}%
4. Stack Overflow. How to create a minimal, reproducible example. 2020.
Available:
\url{https://stackoverflow.com/help/minimal-reproducible-example}

\leavevmode\hypertarget{ref-Bockeler_2020}{}%
5. Böckeler B, Siessegger N. On pair programming. 2020. Available:
\url{https://martinfowler.com/articles/on-pair-programming.html}

\leavevmode\hypertarget{ref-Wickham_2016}{}%
6. Wickham H, Grolemund G. O'Reilly Media; 2016. Available:
\url{https://r4ds.had.co.nz}

\leavevmode\hypertarget{ref-Schloss_2018}{}%
7. Schloss PD. The riffomonas reproducible research tutorial series.
Journal of Open Source Education. 2018;1: 13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}

\leavevmode\hypertarget{ref-Noble2009}{}%
8. Noble WS. A quick guide to organizing computational biology projects.
Lewitter F, editor. PLoS Computational Biology. 2009;5: e1000424.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000424}{10.1371/journal.pcbi.1000424}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.25\columnwidth}\raggedright
\textbf{Title}\strut
\end{minipage} & \begin{minipage}[b]{0.69\columnwidth}\raggedright
\textbf{Description}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.25\columnwidth}\raggedright
base vs.~ggplot\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given input data and a figure, recreate the figure using R's base
graphics or ggplot syntax\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Snakemake\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a bash script that contains an analysis pipeline, convert it to a
Snakemake workflow (can also be done with GNU Make)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
DRYing code\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given script with repeated code, create functions to remove
repetition\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
mothur and Vegan\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a pairwise community dissimilarity matrix, compare communities
using the \texttt{adonis} function in the Vegan R package\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
tidy data\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a wide-formatted data table, convert it to a long, tidy-formatted
data table using tools from R's tidyverse\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
GitFlow\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Participants file and claim an issue to add their name to a README file
in a GitHub-hosted repository and file a pull request to complete the
issue\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
R with Google docs\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Scrape a Google docs workbook and clean the data to identify previous
Code Club presenters\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Develop an R Package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Convert a lab member's collection of scripts into package over a series
of sessions\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

>>>>

\vspace{5mm}

Ada K. Hagan\({^{1,2}}\), Nicholas A. Lesniak\({^1}\), Marcy J.
Balunas\({^3}\), Lucas Bishop\({^1}\), William L. Close\({^1}\), Matthew
D. Doherty\({^1}\), Amanda G. Elmore\({^1}\), Kaitlin J.
Flynn\({^{1,4}}\), Geoffrey D. Hannigan\({^{1,5}}\), Charlie C.
Koumpouras\({^{1,6}}\), Matthew L. Jenior\({^{1,7}}\), Ariangela J.
Kozik\({^8}\), Kathryn McBride\({^1}\), Samara B. Rifkin\({^8}\), Joshua
M.A.~Stough\({^1}\), Kelly L. Sovacool\({^9}\), Marc A.
Sze\({^{1,10}}\), Sarah Tomkovich\({^1}\), Begum D.
Topcuoglu\({^{1,5}}\), and Patrick D. Schloss\({^1}\)\({^\dagger}\)

\vspace{5mm}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Department of Microbiology and Immunology, University of Michigan, Ann
  Arbor, MI
\item
  Current affiliation: Alliance SciComm \& Consulting, Linden, MI
\item
  Division of Medicinal Chemistry, Department of Pharmaceutical
  Sciences, University of Connecticut, Storrs, CT
\item
  Current affiliation: Benaroya Research Institute, Seattle, WA
\item
  Current affiliation: Exploratory Science Center, Merck \& Co., Inc.,
  Cambridge, Massachusetts, USA
\item
  Current affiliation: Roche Diagnostics, Clinical Operations Services
  \& eSystems, Indianapolis, IN
\item
  Current affiliation: Department of Biomedical Engineering, University
  of Virginia, Charlottesville, VA
\item
  Department of Internal Medicine, University of Michigan
\item
  Department of Computational Medicine \& Bioinformatics, University of
  Michigan
\item
  Current affiliation: Genomics and Pharmacogenomics, Merck \& Co.,
  Inc., Cambridge, MA
\end{enumerate}

\vspace{5mm}

@@

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing data
generation and data analysis, we have found scientists feel empowered
when they can both ask and answer their own biological questions
{[}1{]}. Yet, the standard undergraduate and graduate training in the
biological sciences is typically insufficient for developing these
advanced data analysis skills. In our experience performing microbiome
research, it is more common to find exceptional bench scientists who are
inexperienced at analyzing large data sets than to find the reverse. Of
course this raises a challenge: How do we train bench scientists to
effectively answer biological questions with these larger datasets?

The ever-growing ability to generate data and feeling of unease in
analyzing the data are analogous to the struggles scientists also face
with engaging the voluminous scientific literature. A common strategy
for keeping up with the literature is participating in journal clubs,
which involve group discussion of a pre-selected paper. These range from
informal discussions to PowerPoint presentations and course credit. In
addition to staying current on the literature, journal clubs help
strengthen skills in critical thinking, communication, and integrating
the literature {[}2{]}. Our research group has leveraged the similarity
between the overwhelming nature of both scientific literature and data
analysis to address the challenge of teaching reproducible data analysis
practices to bench scientists. Over the past four years we have
experimented with a Code Club model to improve data analysis skills in a
community environment.

Our Code Club sessions are an hour long and alternate with Journal Club
as the second part of weekly two hour lab meetings. Initially, the Code
Club was used to review code from trainee projects. Instead of a
presentation, the presenter would project their code onto a screen and
the participants would go through the code, stating the logic behind
each line. This approach emphasized the importance of code readability
and gave beginners the opportunity to see the real-life, messy code of
more experienced peers. Unfortunately, the format only allowed us to
review a fraction of a project's code, making it difficult to integrate
the programmer's logic across their full project. A major issue with
this model was that sometimes beginners could not contribute to
improving the code, and even when they could, more experienced group
members would eventually dominate the discussion. This led to a lack of
participation by beginners, who would sometimes mentally check out and
resulted in an adversarial environment between the presenter and more
experienced members. As a result, presenters were reluctant to offer to
present again.

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}3{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using either R syntax from the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. Our Code Clubs
typically have 7 to 10 participants, but the inherent
``think-pair-share'' strategy should allow it to be scaled to larger
groups {[}4{]}.

We continue to experiment with approaches for running Code Club. During
the period of the Covid-19 pandemic in 2020 when many research labs were
shuttered, PDS started hosting an open virtual weekly Code Club.
Participants could engage the material posted to
\url{https://www.riffomonas.org/code_club/} via a Zoom meeting or
asynchronously using the recorded sessions available on YouTube.
Regardless of our approach to leading a Code Club, we have learned that
it is critical for the presenter to clearly articulate their goals and
facilitate participant engagement. Although some Code Club sessions may
be more experimental than others, on the whole they are a critical tool
to train bench scientists in reproducible data analysis practices. We
have provided some examples of successful topics in Table 1. We have
summarized the results of our own experimentation as Code Club
presenters and participants into Ten Rules. The first 3 rules apply to
all participants, Rules 4 through 8 are targeted to presenters, and the
last two for participants.

@@

It is critical that the presenter and participants respect each other
and that a designated individual (e.g.~the lab director) enforces a code
of conduct. Each member of the Code Club must have the humility to
acknowledge that they have more to learn about any given topic. We have
found that many problems are avoided when the presenter takes charge of
the session with a clear lesson plan, thoughtfully creates groups, and
gives encouragement. Similarly, participants foster a positive
environment by remembering that the task is not a competition, instead
focusing on the presenter's goals, allowing their partner to contribute,
asking clarifying questions when appropriate, and avoiding distractions
(e.g.~email, social media). Learning to program is challenging. Too
often it is attempted in a environment with nonconstructive criticism.
All parties in a Code Club are responsible for preventing this by
demonstrating respect for themselves and their colleagues.

@@

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it often led to frustration for both
the presenter and participants. If the presenter shared their own code,
did they want participants to focus on their coding style or did they
want help incorporating a new package into their workflow? Participants
will always notice or ask about code concepts that are not the focus of
the exercise; a presenter with a specific goal can bring tangential
conversations back to the planned task. Where possible, presenters
should create a simplified scenario (i.e.~minimal, reproducible example
{[}5{]}), which can be helpful in focusing the participants. The
presenter should verify ahead of time that the simplified example works
and behaves the way they expect. Beyond the content, clear goals for
participant activities will help both parties stay on task and avoid
frustration. For more advanced learners, the presenter can create
stretch goals or give an activity with multiple stopping points where
participants would feel successful (e.g.~commenting code, creating
function, implementing function, refactoring function). Accomplishing
specific goals is more likely to result in a positive outcome for
presenters and participants.

@@

Participants want to learn topics that will either be useful to them or
help their colleague (i.e.~the presenter). Presenters should do their
best to satisfy those motivations, whether it is through the relevancy
of the concept or the data. It does not make sense to present a Code
Club on downloading stock market data if it is not useful or interesting
to the group. Similarly, participants should not be tasked with
improving the presenter's code if the presenter has no intention of
incorporating the suggestions. A list of packages or tasks that group
members are interested in could help a presenter struggling to find a
topic choose one that will make a rewarding Code Club.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. In these circumstances it can be useful to cover commonly
used tools but that might result in disinterest of more experienced
participants. We have identified several strategies to overcome the
challenges presented by participants at varying skill levels. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on their goals. Partnerships between those with similar skill
levels requires the presenter to design appropriate activities for each
skill level. We have found that commenting code is a good skill for
beginners since it forces them to dissect and understand a code chunk
line-by-line. It also reinforces the value of commenting as they develop
their skills and independence. An advantage of forming partnership
between people with disparate skill levels is that it is more likely for
groups to provide the presenter with a diverse range of methods that
achieve the same result. This approach to pairing also helps to graft
new members that have emerging programming skills. Regardless of how
partners are selected, consider asking the pairs to identify a navigator
and a driver {[}6{]}. The driver types at the computer while the
navigator tells them what to type, thus ensuring participation of both
partners. Midway through the activity, the presenter can have the
partners switch roles. Intentionally forming pairs can also engineer
group interactions by avoiding potentially disruptive partnerships or
pairing reliable role models with new group members.

@@

It is not possible to fully participate in a Journal Club discussion
about a paper that the participant has not read. In that context, coming
to Code Club without having installed a dependency is similar to asking
a simple question about the Journal Club paper. Both instances show a
lack of preparation. Just as a presenter must follow Rule 6 to provide
materials ahead of time, participants must review the code in advance,
download the data sets, install the necessary packages, and perhaps read
up on the topic. If the Code Club is based on a paper or chapters in a
data science book (e.g., {[}7--9{]}), the participants should read them
before the session and consider how they might incorporate the concepts
into their own work.

@@

An essential ingredient of any Code Club is active participation from
all parties. Having an open laptop on the table and permission to use it
can feel like an invitation to get distracted by other work, emails, and
browsing the internet; fight that urge and focus on the presenter's
goals. Be respectful and allow your partner to contribute. Speak up for
yourself and force your partner to let you contribute. If the material
seems too advanced for you, it can be frustrating, and tempting to
mentally check out. Fortunately, programming languages like R and Python
are generally expressive, which should allow you to engage with the
logic, even if the syntax is too advanced. Oftentimes, understanding the
logic of \emph{when} to use one modeling approach over another is more
important than knowing \emph{how} to use it. If you understand the
``why'', the ``how'' will quickly follow. More experienced participants
should aim to communicate feedback and coding suggestions at a level
that all participants can understand and engage in. Regardless of skill
levels, your partner and the presenter put themselves in a vulnerable
position by revealing what they do or do not know. Encourage them, and
show your gratitude for helping you learn something new, by fully
participating in each Code Club.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Carey2018}{}%
1. Carey MA, Papin JA. Ten simple rules for biologists learning to
program. Markel S, editor. PLOS Computational Biology. 2018;14:
e1005871.
doi:\href{https://doi.org/10.1371/journal.pcbi.1005871}{10.1371/journal.pcbi.1005871}

\leavevmode\hypertarget{ref-Lonsdale2016}{}%
2. Lonsdale A, Penington JS, Rice T, Walker M, Dashnow H. Ten simple
rules for a bioinformatics journal club. Lewitter F, editor. PLOS
Computational Biology. 2016;12: e1004526.
doi:\href{https://doi.org/10.1371/journal.pcbi.1004526}{10.1371/journal.pcbi.1004526}

\leavevmode\hypertarget{ref-Wilson2014}{}%
3. Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et al.
Best practices for scientific computing. Eisen JA, editor. PLoS Biology.
2014;12: e1001745.
doi:\href{https://doi.org/10.1371/journal.pbio.1001745}{10.1371/journal.pbio.1001745}

\leavevmode\hypertarget{ref-Kothiyal2014}{}%
4. Kothiyal A, Murthy S, Iyer S. Think-pair-share in a large CS1 class.
Proceedings of the 2014 conference on innovation \& technology in
computer science education - ITiCSE 14. ACM Press; 2014.
doi:\href{https://doi.org/10.1145/2591708.2591739}{10.1145/2591708.2591739}

\leavevmode\hypertarget{ref-StackOverflow_MREE}{}%
5. Stack Overflow. How to create a minimal, reproducible example. 2020.
Available:
\url{https://stackoverflow.com/help/minimal-reproducible-example}

\leavevmode\hypertarget{ref-Bockeler_2020}{}%
6. Böckeler B, Siessegger N. On pair programming. 2020. Available:
\url{https://martinfowler.com/articles/on-pair-programming.html}

\leavevmode\hypertarget{ref-Wickham_2016}{}%
7. Wickham H, Grolemund G. O'Reilly Media; 2016. Available:
\url{https://r4ds.had.co.nz}

\leavevmode\hypertarget{ref-Schloss_2018}{}%
8. Schloss PD. The riffomonas reproducible research tutorial series.
Journal of Open Source Education. 2018;1: 13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}

\leavevmode\hypertarget{ref-Noble2009}{}%
9. Noble WS. A quick guide to organizing computational biology projects.
Lewitter F, editor. PLoS Computational Biology. 2009;5: e1000424.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000424}{10.1371/journal.pcbi.1000424}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.25\columnwidth}\raggedright
\textbf{Title}\strut
\end{minipage} & \begin{minipage}[b]{0.69\columnwidth}\raggedright
\textbf{Description}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.25\columnwidth}\raggedright
base vs.~ggplot2\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given input data and a figure, recreate the figure using R's base
graphics or ggplot2 syntax\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Snakemake\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a bash script that contains an analysis pipeline, convert it to a
Snakemake workflow (can also be done with GNU Make)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
DRYing code\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given script with repeated code, create functions to remove
repetition\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
mothur and Vegan\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a pairwise community dissimilarity matrix, compare communities
using the \texttt{adonis} function in the Vegan R package\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
tidy data\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a wide-formatted data table, convert it to a long, tidy-formatted
data table using tools from R's tidyverse\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
GitFlow\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Participants file and claim an issue to add their name to a README file
in a GitHub-hosted repository and file a pull request to complete the
issue\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
R with Google docs\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Scrape a Google docs workbook and clean the data to identify previous
Code Club presenters\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Develop an R Package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Convert a lab member's collection of scripts into an R package over a
series of sessions\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}
